Slitasje og kjørespor

### 2022-11-30 DRAFT_slitasjeOgKjørespor.Rmd
combine ano and naturetypes

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

@@

Where we have data, this is the distribution of values:
```{r, warning=F}
ggplot(ano, aes(x = andel_hovedoekosystem_punkt))+
         geom_histogram(fill = "grey",
                        colour="black",
                        binwidth = 1)+
  theme_bw(base_size = 12)+
  xlab("Percentage cover of the Våtmark main ecosystem\n in the 250m2 circle")+
  scale_x_continuous(limits = c(0,101),
                     breaks = seq(0,100,10))
```

We can see that people tend to record the variable in steps of 5, and that most sample points are 100% belonging to the same main ecosystem.

We want to use area weighting in this indicator, so we can use this percentage cover data to calculate the area.
```{r}
ano$area_of_våtmark <- (ano$andel_hovedoekosystem_punkt/100)*250
```

Let's now look at the distribution of the variables. First I need to slip the cell values to make them numeric.

```{r}
ano$'7SE' <- as.numeric(sub(pattern = "7SE_",
                 replacement = "",
                 x = ano$bv_7se))
```

The NA's is the case of blank cells. The field app should not allow users leaving this field blank. I will need to remove these rows.
```{r}
ano <- ano[!is.na(ano$`7SE`),]
```

```{r}
ano$'7TK' <- as.numeric(sub(pattern = "7TK_",
                 replacement = "",
                 x = ano$bv_7tk))
```

```{r}
par(mfrow=c(1,2))
barplot(table(ano$`7TK`), xlab="7TK scores")
barplot(table(ano$`7SE`), xlab="7SE scores")
```

### Combine datasets
We need to combine the nature type dataset with th ano dataset

```{r}
naturetypes
```

#### Vegetation zones (bioregions)
I want to look at if the indicator values can be averages over smaller and more meaningfu areas (ecological homogeneous areas) than the regions.
```{r}

path <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp",
               "/data/R/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp")
```

>>>>

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
library(units)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

@@

Where we have data, this is the distribution of values:
```{r, warning=F, fig.cap="Percentage cover of the Våtmark main ecosystem in the 250m2 circle"}
ggplot(ano, aes(x = andel_hovedoekosystem_punkt))+
         geom_histogram(fill = "grey",
                        colour="black",
                        binwidth = 1)+
  theme_bw(base_size = 12)+
  xlab("Percentage cover of the Våtmark main ecosystem\n in the 250m2 circle")+
  scale_x_continuous(limits = c(0,101),
                     breaks = seq(0,100,10))
```

We can see that people tend to record the variable in steps of 5%, and that most sample points are 100% belonging to the same main ecosystem.

We want to use area weighting in this indicator, so we can use this percentage cover data to calculate the area.
Note that both datasets use m2 as area units.
```{r}
ano$area <- (ano$andel_hovedoekosystem_punkt/100)*250
```

Let's now look at the distribution of the variables. First I need to slip the cell values to make them numeric.
```{r}
ano$var_7SE <- as.numeric(sub(pattern = "7SE_",
                 replacement = "",
                 x = ano$bv_7se))
```

The NA's is the case of blank cells. The field app should not allow users leaving this field blank. I will need to remove these rows.
```{r}
ano <- ano[!is.na(ano$var_7SE),]
```

```{r}
ano$var_7TK <- as.numeric(sub(pattern = "7TK_",
                 replacement = "",
                 x = ano$bv_7tk))
```

```{r, fig.cap="Distribution of 7TK and 7SE scores in the ANO data"}
par(mfrow=c(1,2))
barplot(table(ano$var_7TK), xlab="7TK scores")
barplot(table(ano$var_7SE), xlab="7SE scores")
```

Combining these two variables into the indicator, same as for the nature type data.
```{r}
temp <- as.data.frame(ano)
class(temp)
temp$slitasje <- apply(temp[,c("var_7TK", "var_7SE")], 1, max, na.rm=T)
ano$slitasje <- temp$slitasje
```

### Combine datasets
We need to combine the nature type dataset with th ano dataset
I will tag the ano rows in the nature type column, and add data abot the main ecosystem
```{r}
ano$naturtype <- "ANO"
ano$hovedøkosystem <- "Våtmark"
ano$kartleggingsår <- ano$aar
```

I will loose the sf class and the crs when joining the datasets, so I will save the crs
```{r}
sf::st_crs(ano) == sf::st_crs(naturetypes) # TRUE
myCRS <- st_crs(ano)
```

Fix class
```{r}
naturetypes$kartleggingsår <- as.numeric(naturetypes$kartleggingsår)
naturetypes$area <- units::drop_units(naturetypes$area)
```

I can reduce the number of column to keep things a bit more tidy
```{r}
temp <- as.data.frame(ano)
temp2 <- as.data.frame(naturetypes)
slitasje_data <- full_join(select(temp,
                                  GlobalID,
                                  kartleggingsår,
                                  hovedøkosystem,
                                  area,
                                  slitasje,
                                  SHAPE), 
                           select(temp2,
                                  identifikasjon_lokalid,
                                  hovedøkosystem,
                                  kartleggingsår,
                                  area,
                                  slitasje,
                                  SHAPE))
# convert back to sf object
slitasje_data <- st_as_sf(slitasje_data)
st_crs(slitasje_data) <- myCRS
```

```{r, fig.cap="Slitasje scores."}
temp <- slitasje_data %>%
  group_by(slitasje) %>%
  summarise(sum = n())

#### Vegetation zones (bioregions)
I want to look at if the indicator values can be averages over smaller and more meaningful areas (ecological homogeneous areas) than the 5 default regions.
```{r}
path <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp",
               "/data/R/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp")
```
